\documentclass[12pt]{article}
\usepackage{amsmath, amssymb}
\usepackage{geometry}
\usepackage{hyperref}
\usepackage{physics}
\usepackage{cite}

\geometry{margin=1in}

\title{Filter Functions for Dynamical Decoupling Sequences}
\author{osipra}
\date{}

\begin{document}

\maketitle

\begin{abstract}
We derive explicit filter functions $F(\omega, t)$ for the Ramsey, Hahn echo, CPMG-$n$, XY-8, and UDD-$n$ pulse sequences. Each derivation follows from direct evaluation of the squared Fourier transform of the modulation function $f(t')$, as established in document 02. We characterize each filter function's spectral sensitivity and discuss its implications for noise spectroscopy.
\end{abstract}

\section{Ramsey Sequence}

For the Ramsey sequence there are no $\pi$ pulses, so $f(t') = 1$ for all $t' \in [0, t]$. The filter function integral evaluates to:

\begin{equation}
    \int_0^t e^{i\omega t'}\, dt' = \frac{1}{i\omega}\left(e^{i\omega t} - 1\right)
\end{equation}

Taking the magnitude squared:

\begin{equation}
    F_\text{Ramsey}(\omega, t) = \frac{1}{\omega^2}\left(e^{i\omega t} - 1\right)\left(e^{-i\omega t} - 1\right) = \frac{2 - 2\cos(\omega t)}{\omega^2}
\end{equation}

Using the identity $1 - \cos\theta = 2\sin^2(\theta/2)$:

\begin{equation}
    \boxed{F_\text{Ramsey}(\omega, t) = \frac{4\sin^2(\omega t/2)}{\omega^2}}
\end{equation}

In the limit $\omega \to 0$, using $\sin(\omega t/2) \approx \omega t/2$:

\begin{equation}
    F_\text{Ramsey}(\omega \to 0, t) = t^2
\end{equation}

The Ramsey filter function peaks at $\omega = 0$ with maximum value $t^2$. It is a \textbf{low-pass filter} — maximally sensitive to slow, DC-like noise and quasistatic fluctuations. This is why Ramsey dephasing in superconducting qubits is dominated by $1/f$ noise \cite{Bylander2011}.

\section{Hahn Echo Sequence}

For the Hahn echo, a single $\pi$ pulse is applied at $t/2$, so:

\begin{equation}
    f(t') = \begin{cases} +1 & 0 \leq t' < t/2 \\ -1 & t/2 \leq t' \leq t \end{cases}
\end{equation}

The filter function integral splits into two parts:

\begin{equation}
    \int_0^t f(t')\, e^{i\omega t'}\, dt' = \int_0^{t/2} e^{i\omega t'}\, dt' - \int_{t/2}^{t} e^{i\omega t'}\, dt'
\end{equation}

Evaluating each integral:

\begin{align}
    \int_0^{t/2} e^{i\omega t'}\, dt' &= \frac{1}{i\omega}\left(e^{i\omega t/2} - 1\right) \\
    \int_{t/2}^{t} e^{i\omega t'}\, dt' &= \frac{1}{i\omega}\left(e^{i\omega t} - e^{i\omega t/2}\right)
\end{align}

Combining:

\begin{equation}
    \frac{1}{i\omega}\left(2e^{i\omega t/2} - 1 - e^{i\omega t}\right)
\end{equation}

Factoring out $e^{i\omega t/2}$ and using $2 - e^{-i\omega t/2} - e^{i\omega t/2} = 2 - 2\cos(\omega t/2) = 4\sin^2(\omega t/4)$:

\begin{equation}
    = \frac{e^{i\omega t/2}}{i\omega}\cdot 4\sin^2\left(\frac{\omega t}{4}\right)
\end{equation}

Taking the magnitude squared (note $|e^{i\omega t/2}|^2 = 1$):

\begin{equation}
    \boxed{F_\text{Hahn}(\omega, t) = \frac{16\sin^4(\omega t/4)}{\omega^2}}
\end{equation}

In the limit $\omega \to 0$:

\begin{equation}
    F_\text{Hahn}(\omega \to 0, t) \propto \omega^2 t^4 \to 0
\end{equation}

The Hahn echo filter function \textbf{vanishes at $\omega = 0$}, confirming that the $\pi$ pulse refocuses slow noise completely. The filter peaks at:

\begin{equation}
    \frac{\omega t}{4} = \frac{\pi}{2} \implies \omega_\text{peak} = \frac{2\pi}{t}
\end{equation}

The Hahn echo is a \textbf{bandpass filter} centered at $\omega = 2\pi/t$. By varying $t$, the filter peak sweeps across the noise spectrum \cite{Cywinski2008}.

\section{CPMG-\texorpdfstring{$n$}{n} Sequence}

CPMG-$n$ consists of $n$ equally spaced $\pi$ pulses at times:

\begin{equation}
    t_j = \frac{(2j-1)t}{2n}, \quad j = 1, 2, \ldots, n
\end{equation}

The modulation function alternates sign at each pulse:

\begin{equation}
    f(t') = (-1)^k, \quad t_k < t' < t_{k+1}
\end{equation}

where $t_0 = 0$ and $t_{n+1} = t$. The filter function integral becomes a telescoping sum:

\begin{equation}
    \int_0^t f(t')\, e^{i\omega t'}\, dt' = \sum_{k=0}^{n} (-1)^k \int_{t_k}^{t_{k+1}} e^{i\omega t'}\, dt' = \sum_{k=0}^{n} (-1)^k \frac{e^{i\omega t_{k+1}} - e^{i\omega t_k}}{i\omega}
\end{equation}

Adjacent terms in the sum share boundary exponentials with opposite signs, leading to significant cancellation. After collecting terms, substituting $t_k = (2k-1)t/2n$, and summing the resulting geometric series, the sum evaluates to a closed form. Taking the magnitude squared gives \cite{Cywinski2008}:

\begin{equation}
    \boxed{F_\text{CPMG-n}(\omega, t) = \frac{16\sin^4\!\left(\dfrac{\omega t}{4n}\right)}{\omega^2} \cdot \frac{\sin^2\!\left(\dfrac{\omega t}{2}\right)}{\sin^2\!\left(\dfrac{\omega t}{2n}\right)}}
\end{equation}

For $n=1$ this reduces exactly to $F_\text{Hahn}(\omega, t) = 16\sin^4(\omega t/4)/\omega^2$, confirming consistency with the Hahn echo result.

The interference factor $\sin^2(\omega t/2)/\sin^2(\omega t/2n)$ is maximized when $\omega t/2n = \pi/2$, giving the approximate peak location:

\begin{equation}
    \omega_\text{peak} \approx \frac{n\pi}{t}
\end{equation}

By varying both $n$ and $t$, the filter peak can be swept across the noise spectrum, enabling reconstruction of $S(\omega)$ \cite{Alvarez2011}. Note that the true numerical peak is shifted slightly below $n\pi/t$ due to the $1/\omega^2$ denominator, consistent with the analogous shift observed for the Hahn echo.

\section{XY-8 Sequence}

XY-8 is a phase-cycled dynamical decoupling sequence consisting of 8 $\pi$ pulses applied along alternating $X$ and $Y$ axes in the pattern X--Y--X--Y--Y--X--Y--X, with equally spaced pulse times identical to CPMG-8:

\begin{equation}
    t_j = \frac{(2j-1)t}{16}, \quad j = 1, 2, \ldots, 8
\end{equation}

For dephasing noise (Z-axis noise), the modulation function $f(t')$ is identical to CPMG-8, since both X and Y rotations flip the sign of $\sigma_z$ phase accumulation equally. Therefore:

\begin{equation}
    \boxed{F_\text{XY-8}(\omega, t) = F_\text{CPMG-8}(\omega, t) = \frac{16\sin^4\!\left(\dfrac{\omega t}{32}\right)}{\omega^2} \cdot \frac{\sin^2\!\left(\dfrac{\omega t}{2}\right)}{\sin^2\!\left(\dfrac{\omega t}{16}\right)}}
\end{equation}

The filter peaks at $\omega_\text{peak} \approx 8\pi/t$.

The advantage of XY-8 over CPMG-8 lies not in the filter function but in \textbf{robustness to pulse errors}. Imperfect $\pi$ pulses leave residual rotations that accumulate coherently in CPMG. The alternating X/Y phase cycling in XY-8 causes these residual errors to cancel to first order, significantly improving performance on real hardware \cite{Gullion1990}.

\subsection*{Note on pulse error suppression}

The filter function formalism assumes ideal $\pi$ pulses. On IBM hardware, gate fidelities are finite and systematic pulse errors are present. For this reason, XY-8 is preferred over CPMG-8 for hardware experiments in this project, even though their filter functions are theoretically identical.

\section{UDD-\texorpdfstring{$n$}{n} Sequence}

Uhrig Dynamical Decoupling (UDD) \cite{Uhrig2007} places $n$ $\pi$ pulses at nonuniform times determined by:

\begin{equation}
    t_j = t \cdot \sin^2\!\left(\frac{j\pi}{2(n+1)}\right), \quad j = 1, 2, \ldots, n
\end{equation}

This specific spacing is derived by requiring that the coherence decay $\chi(t)$ vanishes to order $n$ in $t$ for a generic smooth $S(\omega)$. Uhrig showed that this is the optimal pulse placement for suppressing dephasing with exactly $n$ pulses \cite{Uhrig2007}, in the sense that no other $n$-pulse sequence can achieve higher-order cancellation.

\subsection*{Filter Function}

The filter function follows the same general sum as CPMG:

\begin{equation}
    \tilde{F}(\omega, t) = \sum_{k=0}^{n} (-1)^k \frac{e^{i\omega t_{k+1}} - e^{i\omega t_k}}{i\omega}
\end{equation}

with $t_0 = 0$, $t_{n+1} = t$, and $F_\text{UDD}(\omega, t) = |\tilde{F}(\omega, t)|^2$.

Unlike CPMG, the nonuniform pulse spacing prevents summation as a geometric series. \textbf{No closed-form expression exists for $F_\text{UDD}(\omega, t)$} — the filter function must be evaluated numerically for each $(\omega, t, n)$.

\subsection*{Comparison with CPMG}

UDD is optimal for \textbf{smooth} $S(\omega)$ — noise spectra without sharp features. In this regime, UDD achieves the same suppression order as CPMG with fewer pulses, making it more efficient.

However, for noise with sharp spectral features — such as Lorentzian peaks from TLS defects — CPMG can outperform UDD because its uniform spacing produces a filter function with better-defined harmonic structure that can be tuned to avoid TLS resonances \cite{Cywinski2008}.

In this project, both sequences are implemented and their reconstructed $S(\omega)$ are compared, with the expectation that they agree for smooth noise and diverge near TLS peaks.

\section{Summary}

\begin{table}[h]
\centering
\begin{tabular}{llll}
\hline
\textbf{Sequence} & \textbf{Filter function} & \textbf{Peak frequency} & \textbf{Character} \\
\hline
Ramsey      & $4\sin^2(\omega t/2)/\omega^2$   & $\omega = 0$       & Low-pass \\
Hahn echo   & $16\sin^4(\omega t/4)/\omega^2$  & $\omega = 2\pi/t$  & Bandpass \\
CPMG-$n$    & $16\sin^4(\omega t/4n)\sin^2(\omega t/2) / [\omega^2\sin^2(\omega t/2n)]$ & $\omega \approx n\pi/t$ & Tunable bandpass \\
XY-8        & Same as CPMG-8                                                            & $\omega \approx 8\pi/t$ & Phase-robust bandpass \\
UDD-$n$     & No closed form                              & Nonuniform         & Optimal for smooth $S(\omega)$ \\
\hline
\end{tabular}
\caption{Filter functions for dynamical decoupling sequences used in this project.}
\end{table}

\begin{thebibliography}{9}

\bibitem{Cywinski2008}
L. Cywi\'{n}ski, R. M. Lutchyn, C. P. Nave, and S. Das Sarma,
\textit{How to enhance dephasing time in superconducting qubits},
Phys. Rev. B \textbf{77}, 174509 (2008).
\href{https://arxiv.org/abs/0807.4926}{arXiv:0807.4926}

\bibitem{Alvarez2011}
G. A. \'{A}lvarez and D. Suter,
\textit{Measuring the Spectrum of Colored Noise by Dynamical Decoupling},
Phys. Rev. Lett. \textbf{107}, 230501 (2011).

\bibitem{Bylander2011}
J. Bylander et al.,
\textit{Noise spectroscopy through dynamical decoupling with a superconducting flux qubit},
Nature Physics \textbf{7}, 565 (2011).

\bibitem{Gullion1990}
T. Gullion, D. B. Baker, and M. S. Conradi,
\textit{New, compensated Carr-Purcell sequences},
J. Magn. Reson. \textbf{89}, 479 (1990).

\bibitem{Uhrig2007}
G. S. Uhrig,
\textit{Keeping a quantum bit alive by optimized $\pi$-pulse sequences},
Phys. Rev. Lett. \textbf{98}, 100504 (2007).

\end{thebibliography}

\end{document}